---
title: "ST 518 HW1"
author: "Cindy Chen"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine : xelatex 
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Conceptual Questions

1. For each of the following, determine whether the response variable numerical or categorical. If the response variable is categorical, is it binary? If it is not binary, list possible categories for the response variable.

    (a) In a survey, college students were asked how many hours per week they spend on the internet.
    
    **numerical variable**
    
    (b) In a survey, college students were asked, "What percentage of the time that you spend on the internet is not for course work?"
    
    **numerical variable**
    
    (c) In a survey, college students were asked, "What is your primary mode of transportation when traveling between campus and home?"
    
    **categorical variable**
    
    (d) In a survey, college students were asked whether or not they live on campus.
    
    **categorical variable**
    
    (e) In a survey, college students were asked how many of their meals they prepare at home per week.
    
    **numerical variable**
    
    (f) In a survey, college students were asked, "Which of the five main food groups constitutes the majority of your diet?"
    
    **categorical variable**
    
    (g) In a survey, smart phone users were asked whether or not they have used a web-based taxi service like Uber or Lyft.
    
    **categorical variable**
    
    (h) In a survey, smart phone users were asked how many times they used a web-based taxi service in the past three months.
    
    **numerical variable**
    
 
 2. On a multiple-choice exam, each of the 20 questions has 2 possible answers and only one correct response. Suppose, for each question, one student selects his responses completely at random.
 
    (a) Let $X_i,\ i = 1,...,20$ represent whether the student answered the $i$th question correctly, and let $Y$ represent the total number of answers the student gets correct. What is the distribution of $X_i$? What is the distribution of $Y$?
    
    **Binomial distribution; Y∼Binomial(n=20,p= 1/2)**
    
    (b) What is the probability that the student passes the test (i.e., scores at least 70\%)? (You will learn more about this distribution in the next module, but you should be able to calculate this in R using the `pbinom` function.)
    
    **Answer: 0.05765915; Solution: 0.7×20=14, 1 - pbinom(13, size = 20, prob = 0.5, lower.tail = TRUE)**
    
 
3. Do you prefer taking courses online, on campus, or a hybrid of the two? Suppose these preferences occur with probabilities ($p_1, p_2, p_3$). For $N = 3$ independent subjects, let the observed frequencies be $(n_1, n_2, n_3)$. That is, we observe $n_1$ subjects who prefer taking courses online, etc.

    (a) Explain how you can determine $n_3$ from knowing $n_1$ and $n_2$.
    
    ** Answer: n3 = N− n1− n2**
    
    (b) List all the possible observations $(n_1, n_2, n_3)$, with $n = 3$.
    
    ** Answer:(0,0,3),(0,1,2),(0,2,1),(0,3,0),(1,0,2),(1,1,1),(1,2,0),(2,0,1),
    (2,1,0),(3,0,0).**
    

## R Question

4. Consider the data in {\bf ex2117} of the {\bf Sleuth3} package.

    (a) In what format are these data (case, tabular, frequency, other)? Please explain.

```{r}
library(Sleuth3)
head(ex2117)
```


    
    **Answer: The data format is Tabular data which is organized into rows and 
    columns. In each row represents a case or observation, and each column 
    represents a variable or attribute. From the above dataset, the data is 
    organized into rows corresponding to individual observations or samples, and 
    columns representing different variables or measurements associated with each 
    observation.**
    
    (b) Create a new R data frame for the data, called **HW1data**. In that data frame, create a `NumBoys` column that represents the number of births that were boys (be sure to round to the nearest integer).

```{r}
library(Sleuth3)

data(ex2117)

HW1data <- data.frame(ex2117)

HW1data$PctBoysDecimal <- HW1data$PctBoys / 100

HW1data$NumBoys <- round(HW1data$PctBoysDecimal * HW1data$Number)

head(HW1data)

```
 
  (c) Use an appropriate R command or commands to change the format to something else (you can pick what).
```{r}
library(Sleuth3)

data(ex2117)

ex2117$Group <- as.character(ex2117$Group)
ex2117$Time <- as.character(ex2117$Time)
ex2117$PctBoys <- paste0(round(ex2117$PctBoys, 0), "%")
head(ex2117)

```