---
title: "ST 518 HW2"
author: "Cindy Chen"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine : xelatex 
    
---

1. (4 points) It is often argued that victims of violence exhibit more violent 
behavior toward others.To study this hypothesis, a researcher searched court 
records to find 908 individuals who had been victims of abuse as children. She 
then found 667 individuals, with similar demographic characteristics, who had not 
been abused as children. Based on a search through subsequent years of court 
records, she was able to determine how many in each of these groups became involved 
in violent crimes, as shown in the following table.
Involved in violent crime?
Yes No
Abuse Victim 102 806
Control 53 614
The researcher concluded: ”Early childhood victimization has demonstrable long-term 
consequences for violent criminal behavior.” Conduct your own analysis of the 
data and comment on this conclusion. Is there evidence of a difference between 
the two groups? Is the strength of the causal implication of this statement
justified by the data from this study?


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

abuse_victim <- c(Yes = 102, No = 806)
control <- c(Yes = 53, No = 614)



```

$$ H0: There \ is \ no \ association \ between \ the \ two \ groups.$$
$$ H1: There \ is \  association \ between \ the \ two \ groups.$$
```{r}

contingency_table <- matrix(c(abuse_victim, control), nrow = 2, byrow = TRUE)
contingency_table
```

```{r}
chi_squared_test <- chisq.test(contingency_table)
chi_squared_test
```


Answer:

Yes, there is evidence of a difference between the two groups. The p-values of 
0.03765 indicates that there is a statistically significant difference 
between the 2 groups (abuse victims and controls)  in terms of their likelihood 
of exhibiting violent criminal behavior.

Therefore, there is evidence to suggest that there is indeed a difference between
the two groups. However, the statistical analysis shows a significant association, 
it does not necessarily imply there is causation.



Conceptual Questions
2. (1 point) During an investigation of the U.S. space shuttle Challenger disaster, 
it was learned that project managers had judged the probability of mission failure 
to be 0.00001, whereas engineers working on the project had estimated failure 
probability at 0.005. The difference between these two probabilities, 0.00499, 
was discounted as being too small to worry about. Is a different picture
provided by considering odds? How is that interpreted?

```{r}
pm_fail <- 0.00001
pm_success <- 1 - 0.00001
eng_fail <- 0.005
eng_success <- 1 -  0.005

```

```{r}
pm_fail_odds <- pm_fail / pm_success
eng_fail_odds <- eng_fail / eng_success
odds_ratio_1 <- pm_fail_odds / eng_fail_odds
odds_ratio_1
```

Answer:
The odds ratio of approximately 0.00199002 suggests that the odds of failure 
estimated by project managers are approximately 0.002 times the odds estimated
by engineers. This indicates a significant discrepancy between the two groups' 
perceptions of the likelihood of failure.


3. (2 points) Suppose that 90% of orange tabby cats are male. Determine if the following statements
are true or false, and explain your reasoning.
(a) The distribution of sample proportions of random samples of size 30 is left skewed.

Answer: 
False. Based on Central Limit Theorem, when the sample size is large enough, the 
distribution of sample proportion follows normal distribution. The sample size 
of 30 is considered large enough for CLT.

(b) Using a sample size that is 4 times as large will reduce the standard error 
of the sample proportion
by one-half.

Answer:
True. The standard error of sample proportion decreases as the
sample size increases. 


(c) The distribution of sample proportions of random samples of size 140 is 
approximately normal.

Answer:
True. Based on Central Limit Theorem, when the sample size is large enough, the 
distribution of sample proportion follows normal distribution. The sample size of 
140 is considered large enough for CLT.


(d) The distribution of sample proportions of random samples of size 280 is 
approximately normal.

Answer:
True. Based on Central Limit Theorem, when the sample size is large enough, the 
distribution of sample proportion follows normal distribution. The sample size of 
280 is considered large enough for CLT.

4. (2 points) A 2010 survey asked 827 randomly sampled registered voters in California, ”Do you
support or do you oppose drilling for oil and natural gas off the coast of California? Or do you not
know enough to say?” Below is the distribution of responses, separated based on whether or not the
respondent graduated from college.
College Grad
Yes No
Support 154 132
Oppose 180 126
Do not know 104 131
Total 438 389
(a) What percents of college graduates and non-college graduates in this sample do not know
enough to have an opinion on drilling for oil and natural gas off the coast of California (i.e.,
report two percent values)?

Answer:
college_not_know: 104/438 = 23.74%
non_college_not_know: 131/389 = 33.67%


(b) Conduct a hypothesis test to determine whether there is evidence that the proportion of college graduates who do not have an opinion on this issue is different from that of non-college
graduates.

$$ H0: The \ proportion \ of \ college \ graduates \ who \ do \ not \ have \ an \ 
opinion \ is \ the \ same \ as \ the \ proportion \ of \ non-college \ graduates \ 
who \ do \ not \ have \ an \ opinion. $$
$$ H1: The \ proportion \ of \ college \ graduates \ who \ do \ not \ have \ an \ 
opinion \ is \ different \ from \ the \ proportion \ of \ non-college \ graduates \ 
who \ do \ not \ have \ an \ opinion. $$

```{r}

college_grad <- c(104, 438 - 104)
non_college_grad <- c(131, 389 - 131)


test_result <- prop.test(x = college_grad, n = college_grad + non_college_grad,
                         alternative = "two.sided")

print(test_result)

```
Answer:
The p-value(0.002045) is less than significance level of 0.05,
we reject the null hypo. Therefore, there is evidence to suggest that the 
proportion of college graduates who do not have an opinion on the issue is 
different from the proportion of non-college graduates who do not have an opinion.



5. (1 point) A study of British male physicians noted that the proportion who died from lung cancer
was 0.0140 per year for cigarette smokers and 0.00010 per year for nonsmokers. Additionally, the
proportion who died from heart disease was 0.00669 for smokers and 0.00413 for nonsmokers. Which
response (lung cancer or heart disease) is more strongly related to cigarette smoking, in terms of the
reduction in deaths that could occur with the absence of smoking?

Answer:
```{r}

lung_cancer_smokers <- 0.0140
lung_cancer_nonsmokers <- 0.00010


heart_disease_smokers <- 0.00669
heart_disease_nonsmokers <- 0.00413


odds_ratio_lung_cancer <- (lung_cancer_smokers / (1 - lung_cancer_smokers)) / (lung_cancer_nonsmokers / (1 - lung_cancer_nonsmokers))
odds_ratio_heart_disease <- (heart_disease_smokers / (1 - heart_disease_smokers))/ (heart_disease_nonsmokers / (1- heart_disease_nonsmokers))


cat("Odds ratio for lung cancer:", odds_ratio_lung_cancer, "\n")
cat("Odds ratio for heart disease:", odds_ratio_heart_disease, "\n")

```
Answer:

An odds ratio of 141.97 for lung cancer suggests that the odds of dying from lung 
cancer are approximately 142 times higher for smokers compared to nonsmokers.
An odds ratio of 1.62 for heart disease suggests that the odds of dying from heart 
disease are approximately 1.62 times higher for smokers compared to nonsmokers.
Based on these results, we can conclude that the odds of dying from lung cancer 
are much more strongly related to cigarette smoking compared to the odds of dying 
from heart disease.