---
title: "HW Assignment 8"
author: "Cindy Chenr"
date: "`r Sys.Date()`"
output: pdf_document
---


R Question
1. (5 points) Consider the crashi data in the VGAM library that we examined in Lab. Sometimes people
will fit multiple linear regression models to data like this, not recognizing that the responses are counts.
(a) Using the Time.Cat and Day.Cat variables that we created in Lab, fit a multiple linear regression
model to the counts—be sure to include the interaction between Time.Cat and Day.Cat just as we
did in the lab. How does the model fit compare to the negative binomial regression model that you
fit in lab? Please explain.
```{r}
#install.packages('VGAM')
```


```{r}
library(VGAM)
library(MASS)
library(dplyr)
library(magrittr)
hour <- rownames(crashi) ## grab the hours
crashi2 <- stack(crashi) ## combine 7 columns of crashes into 1
names(crashi2) <- c("Count","Day")
crashi2$Day <- factor(crashi2$Day,levels(crashi2$Day)[c(2,6,7,5,1,3,4)])  # make sure the days are ordered correctly
crashi2$Hour <- as.numeric(rep(hour, ncol(crashi)))  
crashi2 %<>% mutate(.,Day.Cat = ifelse((Day != "Fri" & Day != "Sat" & Day != "Sun"),"Weekday",as.character(Day)), Time.Cat = cut(Hour,
	breaks = c(-1, 5.5, 11.5, 18.5, 25),
	labels = c("Early.Morn", "Morn", "Afternoon", "Evening")))
head(crashi2)

## add a column with hour and make it numeric (not categorical)
head(crashi2)
data(crashi2)
head(crashi2)
```
```{r}
unique_days <- unique(crashi2$Day.Cat)
print(unique_days)
```


```{r}
library(VGAM)
lm_model <- lm(Count ~ Time.Cat * Day.Cat, data = crashi2)
summary(lm_model)
```
```{r}
mod3 <- glm.nb(Count ~ Day.Cat * Time.Cat, data = crashi2)
summary(mod3)
```
Answer:
Multiple linear regression model assumes a linear relationship between predictors
and response variable. However, it didn't count the nature of the response variable.
So it may lead to violations of assumptions such as normality and homoscedasticity.
From the Adjusted R-squared value is 0.7402, indicating that approximately 74.02% 
of the variance in the response variable is explained by the predictors. The 
p-values for the coefficients suggest that all predictors are statistically significant.

The negative binomial regression model is specifically designed for count data 
and it will account for overdispersion and the count nature of the response variable.
The dispersion parameter (Theta) is estimated to be approximately 10.55 which 
indicates overdispersion. The model's log-likelihood value is -1405.533 .
The p-values for the coefficients suggest that all predictors are statistically significant.

Therefore, the multiple linear regression model assumes a linear relationship 
between predictors and the response and it may not accurately capture the 
relationship for count data. However, the negative binomial regression model 
provides valid estimates and inferences by accounting for overdispersion and the 
count nature of the response variable. The negative binomial regression model is 
expected to provide a better fit and more reliable estimates for count data 
compared to the multiple linear regression model.



(b) Now compare the predicted number of crashes in the early morning on Saturdays using your model
from part (a) and the negative binomial model from the Lab. Are they the same? 
Different?

```{r}

newdat <- data.frame(Day.Cat = "Sat", Time.Cat = "Early.Morn")


predict(lm_model, newdata = newdat, type = "response")
predict(mod3, newdata = newdat, type = "response")

```
Answer:
The predicted value of both models are identical.

(c) Now examine the standard errors associated with the predictions from part (b). Are they the same?
Different? Which model do you prefer? Please explain.

```{r}
lm_pred <- predict(lm_model, newdata = newdat, type = "response", se.fit = TRUE)
mod3_pred <- predict(mod3, newdata = newdat, type = "response", se.fit = TRUE)


lm_std_err <- lm_pred$se.fit
mod3_std_err <- mod3_pred$se.fit

cat("Standard errors for lm_model:", lm_std_err, "\n")
cat("Standard errors for mod3:", mod3_std_err, "\n")
```
Answer:
The standard errors for the predictions from the two models are different. The 
standard error for lm_model is 8.528403, mod3 it is 4.868124.

The standard error provides a measure of the uncertainty associated with the 
predictions. A lower standard error suggests that the predictions are more 
precise and have less variability. Therefore, from the standard errors, mod3 
would be preferred over lm_model for making predictions.

However, it's important to consider other factors such as model performance, 
interpretability, and the context of the problem before making a final decision 
on which model to prefer.

Conceptual Question
2. (5 points) Please find an article or web posting in which the authors discuss the statistical issues with
large datasets. Write a short paragraph in which you identify at least one of the issues that the authors
raise. Also please comment on whether you agree with or disagree with their assessment. Please submit
either a PDF copy of the article you read, or a web link to the article or post.

Answer:
https://academic.oup.com/nsr/article/1/2/293/1397586

The article mentioned various statistical challenges posed by the increasing 
volume and complexity of data in modern research and industry settings. One of 
the important issues is that the problem of noise accumulation can lead to 
spurious correlations and affect the reliability of statistical inference in
high-dimensional datasets. In this article, it emphasizes the importance of 
developing robust statistical methods to mitigate the impact of noise 
accumulation and ensure the validity of conclusions drawn from large datasets.
I agree with the author's assessment of the issue of noise accumulation in 
high-dimensional datasets. As datasets grow in size and becomes dimensionality, 
it has higher probability to have more spurious correlations and misleading results
as noise increases. However, Traditional statistical methods may not be 
sufficient to address these challenges. Therefore, researchers can ensure the 
integrity and reliability of their analyses to address the issues.

